---
title: "Burnout"
author: "Mahmoud Elsherif"
date: "2026-03-29"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Setup for both analyses

Load packages:

```{r message = FALSE, warning = FALSE}

library(mice) # : Stands for Multivariate Imputation by Chained Equations. It is used to fill in (impute) missing data
library(rjags) #Interface to the JAGS MCMC software. It allows you to run Bayesian models, which is useful for the JointAI package.)
library(JointAI) #Used for Joint Analysis and Imputation. it performs Bayesian analysis of models (like linear mixed models) when you have missing data.
library(VIM) #Used for the Visualization and Imputation of Missing Values. It helps you see where data is missing in your dataset
library(naniar) #A modern Tidyverse-friendly package for visualizing and handling missing data. It does not provide Variance Inflation Factors (VIF)
library(e1071) # Calculate skewness/kurtosis.
library(ordinal) # For logistic regression
library(ggpubr) #To make graphs on multiple panels
library(reshape2)#package for tidyverse
library(rlang) # package for tidyverse
library(broom) #provide tidy model outputs
library(afex) #for p-values
library(viridis)#for colourblind friendly images
library(gt) #for tables
library(tidyverse)

library(here)
data_path <- here("data")
```

Load data:

```{r message = FALSE, warning = FALSE}
# Compound data:

data = read_csv(here::here("data/processed/aggregated_data.csv"))
```

Converts strings and ordinal into numbers

```{r setup, include = FALSE}

data <- data %>%
  mutate(title_num = case_when(
    title == "PhD student" ~ 1,
    title == "Post Doc" ~ 2,
    title == "Assistant Professor" ~ 3,
    title == "Associate Professor" ~ 4,
    title == "Full Professor" ~ 5,
    TRUE ~ 0  # Covers "This does not apply to me" or any others
  ))

data <- data %>%
  mutate(gender_num = case_when(
    gender == "Man" ~ 1,
    gender == "Woman" ~ 2,
    gender == "Non-binary / third gender" ~ 3,
    gender == "Prefer not to say" ~ 4,
    TRUE ~ NA_real_
  ))

data <- data %>%
  mutate(discipline_num = case_when(
    discipline == "Non-STEM" ~ 0,
    discipline == "STEM (science, technology, engineering, mathematics)" ~ 1, # Make sure the text matches exactly
    TRUE ~ NA_real_
  ))

data <- data %>%
  mutate(last_published_num = case_when(
    last_published == "This year" ~ 1,
    last_published == "In the last year" ~ 2,
    last_published == "In the last 5 years" ~ 3,
    last_published == "In the last 10 years" ~ 4,
    TRUE ~ NA_real_
  ))

data <- data %>%
  mutate(academia_length_num = case_when(
    academia_length == "Less than 1 year" ~ 1,
    academia_length == "1-5 years" ~ 2,
    academia_length == "6-10 years" ~ 3,
    academia_length == "More than 10 years" ~ 4,
    TRUE ~ NA_real_
  ))

data <- data %>%
  mutate(phd_dissertation_num = case_when(
    phd_dissertation == "1-2 articles" ~ 1,
    phd_dissertation == "3-4 articles" ~ 2,
    phd_dissertation == "More than 4 articles" ~ 3,
    TRUE ~ 0  # Catch-all for NA and "I don't know"
  ))

data <- data %>%
  select(#-gender, 
         #-title, 
         #-discipline, 
         #-last_published, 
         #-academia_length, 
         #-phd_dissertation, 
         -anonym_id)

```

# Assessing missingness in the data ---------------------------------------

```{r setup, include=FALSE}

# Replace "NA" with proper NA in all columns
data <- data %>%
  mutate(across(where(is.character), ~na_if(., "NA")))

data %>% select(where(is.character)) %>% colnames()

#Assess the missing data patterns
mcar_test(data %>%  select(brt_1:gnl_2)
)

#The pattern is significant, meaning the data is not completely missing at random, thus we will need to use multiple imputation

  ## Missing data patterns
  summary(data)
  md.pattern(data, rotate.names = TRUE)

  # Visual inspection of missing data patterns on item level
  # Plotting the variables
  
  par(mar = c(2,3,2,1), mgp = c(2, 0.6, 0))
  
  
plot_data <- data %>%
  select(brt_1:gnl_2) %>%
  # This 'unclasses' the columns to ensure they only have ONE class (numeric)
  mutate(across(everything(), ~as.numeric(as.character(.))))

final_plot_data <- plot_data %>%
  as.matrix() %>%      # Converting to matrix strips all 'attributes' and extra classes
  as.data.frame()      # Converting back makes it a clean data frame again

# Plot just the 'brt' scale
plot_all(final_plot_data %>% select(starts_with("brt")), nclass = 10)

# Plot just the 'pps' scale
plot_all(final_plot_data %>% select(starts_with("pps")), nclass = 10)

plot_all(final_plot_data %>% select(starts_with("ppa")), nclass = 10)

plot_all(final_plot_data %>% select(starts_with("ppr")), nclass = 10)

plot_all(final_plot_data %>% select(starts_with("sec")), nclass = 10)

plot_all(final_plot_data %>% select(starts_with("wlb")), nclass = 10)


plot_all(final_plot_data %>% select(starts_with("wld")), nclass = 10)

plot_all(final_plot_data %>% select(starts_with("scs")), nclass = 10)

plot_all(final_plot_data %>% select(starts_with("scc")), nclass = 10)

plot_all(final_plot_data %>% select(starts_with("sec")), nclass = 10)

plot_all(final_plot_data %>% select(starts_with("opp")), nclass = 10)

plot_all(final_plot_data %>% select(starts_with("qrp")), nclass = 10)

plot_all(final_plot_data %>% select(starts_with("inf")), nclass = 10)

plot_all(final_plot_data %>% select(starts_with("mng")), nclass = 10)

plot_all(final_plot_data %>% select(starts_with("rla")), nclass = 10)

plot_all(final_plot_data %>% select(starts_with("rlc")), nclass = 10)

plot_all(final_plot_data %>% select(starts_with("pay")), nclass = 10)

plot_all(final_plot_data %>% select(starts_with("tol")), nclass = 10)

plot_all(final_plot_data %>% select(starts_with("gnl")), nclass = 10)

  aggr_plot <- aggr(final_plot_data, col=c('navyblue','red'),
                  numbers=TRUE, sortVars=TRUE, 
                  labels=TRUE, cex.axis=.5, gap=2, 
                  ylab=c("Histogram of missing data","Pattern"))
 
  vis_miss(final_plot_data, sort_miss = TRUE)
  gg_miss_upset(final_plot_data, nsets = n_var_miss(final_plot_data), nintersects = NA)
  gg_miss_case(final_plot_data)
  gg_miss_case_cumsum(final_plot_data)
  gg_miss_var_cumsum(final_plot_data) + labs(x = "Variable", y = "Cumulative sum of missing values")
  
  # Mean,  sd, skewness and kurtosis of variables
  
# Summary table for all survey items
summary_table <- final_plot_data %>%
  summarise(across(brt_1:gnl_2, 
                   list(mean = ~mean(., na.rm=TRUE), 
                        sd   = ~sd(., na.rm=TRUE),
                        skew = ~e1071::skewness(., na.rm=TRUE, type = 2),
                        kurt = ~e1071::kurtosis(., na.rm=TRUE, type = 2))))




DATA_1 <- data %>%
  mutate(across(
    brt_1:gnl_2,
    ~ ifelse(is.na(.), 1, 0),
    .names = "{.col}_dum"
  ))



```

# Correlations for continuous auxiliary variables

```{r setup, include=FALSE}
  
  # 1. Define your dummy variables and your predictors

# This gets all the 0/1 dummy columns you created
dummies <- names(DATA_1 %>% select(ends_with("_dum")
                                   )
                 )

# This gets all the original numeric columns
predictors <- names(DATA_1 %>% select(brt_1:gnl_2)
                    )
                                      
# 2. Run the correlation matrix for just these groups
cor_matrix <- cor(DATA_1[dummies], DATA_1[predictors], use = "pairwise.complete.obs")


# 1. Identify which dummy variables actually have missing data 
# (where the sum of the dummy is > 0)
active_dummies <- DATA_1 %>%
  select(ends_with("_dum")) %>%
  select(where(~ sum(.) > 0)) %>%
  names()

# 2. Run the correlation ONLY for those that have missing values
# This stops the "Standard Deviation is Zero" warning
cor_matrix <- cor(DATA_1[active_dummies], DATA_1[predictors], use = "pairwise.complete.obs")

# 3. View the result
round(cor_matrix, 2)

# 3. View it as a nice table
round(cor_matrix, 2) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Variable") %>%
  gt()

#When you look at this matrix, you are looking for the "Missingness Mechanism":MCAR (Missing Completely at Random): 
#All correlations are near zero. The missingness is just "bad luck" and doesn't depend on your data.
#MAR (Missing at Random): You see significant correlations (e.g., r = .15 ). This means people with higher rla4 were more likely to miss sec_1_dum. This is good to know because you can use those predictors in your mice imputation to fix the bias.
#MNAR (Missing Not at Random): The most difficult type. This happens if the reason for missingness is the variable itself (e.g., people with low VAR1 scores specifically chose not to do VAR2).

```

The correlation matrix between missingness indicators (dummies) and observed scores showed no substantial systematic patterns (r \< .30 for all pairs), supporting the assumption that data are Missing Completely at Random (MCAR)

# Fisher exact test for binary auxiliary variables

```{r setup, include=FALSE}

# 1. Define your lists
dummies <- names(DATA_1 %>% select(ends_with("_dum")
                                   )
)
categories <- c("gender",	"title",	"discipline",	"last_published",	"academia_length",	"phd_dissertation")

# 2. Use a nested loop to create a dataframe of results
# Ensure demographic categories are treated as factors for testing
DATA_CLEAN <- DATA_1 %>%
  mutate(across(c("gender", "title", "discipline", "last_published", "academia_length", "phd_dissertation"), as.factor))

# 1. Identify only those dummies that actually HAVE missing values
active_dummies <- names(DATA_CLEAN %>% 
                          select(ends_with("_dum")) %>% 
                          select(where(~ sum(.) > 0)))

# 2. Run the loop with an extra-safe level check
results <- expand.grid(Dummy = active_dummies, Category = categories) %>%
  rowwise() %>%
  mutate(p_value = {
    # Extract data and remove rows where either is NA
    valid_data <- data.frame(d = DATA_CLEAN[[Dummy]], c = DATA_CLEAN[[Category]]) %>%
      na.omit()
    
    # Check if we still have at least 2 levels in BOTH columns after dropping NAs
    if(length(unique(valid_data$d)) > 1 && length(unique(valid_data$c)) > 1) {
      # Use the cleaned, paired data
      fisher.test(valid_data$d, valid_data$c, simulate.p.value = TRUE, B = 10000)$p.value
    } else {
      NA_real_
    }
  }) %>%
  ungroup()

# 3. View significant results
significant_results <- results %>% filter(p_value < 0.05)
print(significant_results)
#By running these tests, you are checking if missingness is dependent on categorical factors. Example: If gender  have a significant p-value, it means the Treatment group and Control group have different rates of missing data.Action Item: If you find significance here, you must include these categorical variables as "auxiliary variables" when you eventually run your mice imputation. This informs the model why the data is missing and helps it make better "guesses" for the missing values. 

#The missingness in the research practice items (brt_3, brt_4, qrp_6-8) is not "random noise." Instead, it is systematically related to professional demographics—specifically Job Title and Length of Time in Academia.

# 1. Define your lists based on your actual data
# 'dummies' are your DVs (The "Is it missing?" flags)
dummies <- names(DATA_CLEAN %>% select(ends_with("_dum")))

# 'aux_vars' are your IVs (The demographics/factors)
aux_vars <- c("gender", "title", "discipline", "last_published", "academia_length")

# 2. Run all models at once
logistic_results <- expand.grid(dv = dummies, iv = aux_vars) %>%
  group_by(dv, iv) %>%
  do({
    # We use the current dv and iv from the row to build the formula
    current_dv <- .$dv
    current_iv <- .$iv
    
    # Ensure data is passed correctly here
    tidy(glm(as.formula(paste(current_dv, "~", current_iv)), 
             family = "binomial", 
             data = DATA_CLEAN), # <--- FIX: Added your dataset name here
         exponentiate = TRUE)
  }) %>% 
  filter(term != "(Intercept)") %>%
  ungroup()

view(logistic_results)

# Filter for significant predictors of missingness
significant_logistic <- logistic_results %>% filter(p.value < 0.05)
print("Significant Logistic Predictors:")
print(significant_logistic)

#If you don't use Multiple Imputation, your "average" score for brt_3, brt4, qrp_6, qrp_7, qrp_8 or qrp_6 will be heavily skewed toward the opinions of senior academics. By including academia_length in your mice model, the algorithm will "fill in" those missing values using the known profiles of other junior researchers, effectively restoring the voice of the newer cohort to your dataset.

```

## Associations with observed values

```{r}
# 1. Define your variable groups
predictors <- names(DATA_1 %>% select(brt_1:gnl_2))

aux_vars <- c("gender_num",	"title_num",	"discipline_num",	"last_published_num",	"academia_length_num",	"phd_dissertation_num")

# 2. Generate the full matrix at once
# use = "pairwise.complete.obs" is better than "complete.obs" 
# because it keeps more data for each specific pair.
full_cor_matrix <- cor(DATA_1[predictors], DATA_1[aux_vars], 
                       use = "pairwise.complete.obs")

# 3. View the results rounded to 2 decimal places
round(full_cor_matrix, 2)

# 4. View the results in a table
round(full_cor_matrix, 2) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Variable") %>%
  gt()

```

# One-way ANOVA for dichotomous and categorical auxiliary variables 

```{r}
# 1. Define your groups
dvs <- predictors

ivs <- aux_vars

# 2. Map the ANOVA function across all pairs
# 2. Map the Linear Regression function across all pairs
regression_results <- expand.grid(dv = dvs, iv = ivs) %>%
  group_by(dv, iv) %>%
  do({
    # Run the linear model
    fit <- lm(as.formula(paste(.$dv, "~", .$iv)), data = DATA_CLEAN)
    
    # Tidy the output to get coefficients and p-values
    tidy(fit) %>% 
      filter(term != "(Intercept)")
  }) %>%
  ungroup()

# 3. Filter for significant predictors
significant_regressions <- regression_results %>% filter(p.value < 0.05)
print(significant_regressions)
```

# Performing multiple imputation ------------------------------------------

```{r}

# 1. Prepare the data: Use the numeric versions of demographics
# This avoids the "sparse category" problem in Random Forest
MI_DATA <- DATA_CLEAN %>%
  select(brt_1:gnl_2, gender_num, title_num, discipline_num, 
         last_published_num, academia_length_num, phd_dissertation_num)

# 2. Initialize mice
init <- mice(MI_DATA, maxit = 0)
pred <- init$predictorMatrix
meth <- init$method

# 3. Set a clean Predictor Matrix
pred[,] <- 0
survey_items <- names(MI_DATA %>% select(brt_1:gnl_2))
numeric_aux <- names(MI_DATA %>% select(ends_with("_num")))

# Demographics predict Survey Items
pred[survey_items, numeric_aux] <- 1

# 4. Handle "Orphaned" Auxiliary Variables
# If an auxiliary variable (like phd_dissertation_num) has NAs, 
# it needs a predictor too.
for (aux in numeric_aux) {
    # Let title and academia_length predict other demographics
    pred[aux, c("title_num", "academia_length_num")] <- 1
}

# 5. Final Cleanup
diag(pred) <- 0
meth[survey_items] <- "rf"
meth[numeric_aux] <- "pmm" # Use Predictive Mean Matching for numeric codes

# 6. Run Imputation
imp_final <- mice(MI_DATA, 
                  method = meth, 
                  predictorMatrix = pred, 
                  m = 20, 
                  maxit = 10, 
                  seed = 2026)

# If this is 0, your model is perfectly healthy!
nrow(imp_final$loggedEvents)

# If it's NOT 0, look at the reasons:
print(imp_final$loggedEvents)
```

# Evaluating multiple imputation ---------------------------------------------------

```{r}

# 1. Plot convergence (the "hairy caterpillars")
plot(imp_final)

# 2. Density plots: Check if imputed values (red) follow the distribution of observed values (blue)
densityplot(imp_final)

# 3. Stripplot: Check for outliers in the imputed data
stripplot(imp_final, pch = 20, cex = 1.2)

# Extract all imputed datasets
# action="long" stacks them; include=FALSE removes the original data with NAs
imp_data_long <- complete(imp_final, action = "long", include = FALSE)

# Save to CSV
write.csv(imp_data_long, "Survey_Imputed_Data_Long.csv", row.names = FALSE)

# Calculate Mean and SD for every survey item across all imputations
pooled_summary <- imp_data_long %>%
  # Group by the imputation number (.imp) and any categories if needed
  group_by(.imp) %>%
  summarise(across(brt_1:gnl_2, list(mean = ~mean(., na.rm=TRUE), sd = ~sd(., na.rm=TRUE)))) %>%
  # Now average those means/sds across all  imputations
  pivot_longer(cols = - .imp, names_to = "variable_stat", values_to = "value") %>%
  separate(variable_stat, into = c("variable", "stat"), sep = "_(?=[^_]+$)") %>%
  group_by(variable, stat) %>%
  summarise(pooled_value = mean(value)) %>%
  pivot_wider(names_from = stat, values_from = pooled_value)

# View your final pooled table
print(pooled_summary)
write.csv(pooled_summary, "Pooled_Descriptive_Statistics.csv")

```
