---
title: "aggregation"
author: "Dominik Dianovics"
format: html
---

## Researcher burnout and questionable research practice pilot data cleaning

### Processed to aggregated data

#Loading packages

```{r}
library(tidyverse)
library(here)
library(psych)
data_path <- here("data")
```

#Loading data

```{r}
processed = read_csv(here::here("data/processed/processed_data.csv"))
```

### Filtering out invalid data

```{r}
processed_filtered = processed %>% 
  filter(finished == T 
         & consent == T 
         & academic_yes_no == T
         & research_yes_no == T
         & last_published != "More than 10 years ago or Never"
         & attention == "Passed"
  )
```

# Saving item-level data

```{r}
write_csv(processed_filtered, here::here("data/processed/item_level_processed.csv"))
```

# Checking missingness

```{r}
missingness_summary = processed_filtered %>%
  dplyr::select(starts_with(c("bat", "ppqr", "copsoq", "wlbm", "jis", "qrp", "role", "pay", "tool", "general", "open_practice"))) %>%
  summarise(across(everything(), ~ mean(is.na(.)) * 100)) %>%
  pivot_longer(everything(), names_to = "item", values_to = "missingness_percentage")

library(mice)
library(naniar)

processed_filtered %>%
  dplyr::select(starts_with(c("bat", "ppqr_stress", "ppqr_attitude", "ppqr_resources", "copsoq_workload", "copsoq_soc_sup", "copsoq_oppor", "wlbm", "jis", "copsoq_infl", "copsoq_meaning", "role_ambiguity", "role_conflict"))) %>%
  naniar::mcar_test()

# Looking at missingness patterns, I assume there has been an error in the survey programming, as the first 29 rows all have no bat_3 and bat_4 responses, after that the same relationship does not appear, thus I reject these cases and rerun the test.

processed_filtered %>%
  dplyr::select(starts_with(c("bat", "ppqr_stress", "ppqr_attitude", "ppqr_resources", "copsoq_workload", "copsoq_soc_sup", "copsoq_oppor", "wlbm", "jis", "copsoq_infl", "copsoq_meaning", "role_ambiguity", "role_conflict"))) %>%
  slice(-c(1:8)) %>%
  rowwise() %>%
  naniar::mcar_test()

# Case-level missingness
processed_filtered %>%
  dplyr::select(starts_with(c("bat", "ppqr_stress", "ppqr_attitude", "ppqr_resources", "copsoq_workload", "copsoq_soc_sup", "copsoq_oppor", "wlbm", "jis", "copsoq_infl", "copsoq_meaning", "role_ambiguity", "role_conflict"))) %>%
  slice(-c(1:8)) %>%
  naniar::miss_case_summary()

# Variable-level missingness
processed_filtered %>%
  dplyr::select(starts_with(c("bat", "ppqr_stress", "ppqr_attitude", "ppqr_resources", "copsoq_workload", "copsoq_soc_sup", "copsoq_oppor", "wlbm", "jis", "copsoq_infl", "copsoq_meaning", "role_ambiguity", "role_conflict"))) %>%
  slice(-c(1:8)) %>%
  naniar::miss_var_summary()

processed_filtered %>%
  dplyr::select(starts_with(c("bat", "ppqr_stress", "ppqr_attitude", "ppqr_resources", "copsoq_workload", "copsoq_soc_sup", "copsoq_oppor", "wlbm", "jis", "copsoq_infl", "copsoq_meaning", "role_ambiguity", "role_conflict"))) %>%
  slice(-c(1:8)) %>%
  naniar::gg_miss_var()

# Percentage of people having at least one missing value

processed_filtered %>%
  dplyr::select(starts_with(c("bat", "ppqr_stress", "ppqr_attitude", "ppqr_resources", "copsoq_workload", "copsoq_soc_sup", "copsoq_oppor", "wlbm", "jis", "copsoq_infl", "copsoq_meaning", "role_ambiguity", "role_conflict"))) %>%
  slice(-c(1:8)) %>%
  summarise(missing_any = sum(if_any(everything(), is.na)))

# This is the percentage of people having at least one missing value, after removing those with more than 50% missing values (0 individuals) and first 8 rows due to error.
7 / 243 * 100

# Total missing values
processed_filtered %>%
  dplyr::select(starts_with(c("bat", "ppqr_stress", "ppqr_attitude", "ppqr_resources", "copsoq_workload", "copsoq_soc_sup", "copsoq_oppor", "wlbm", "jis", "copsoq_infl", "copsoq_meaning", "role_ambiguity", "role_conflict"))) %>%
  slice(-c(1:8)) %>%
  summarise(total_missing = sum(is.na(.)),
            total_values = nrow(.) * ncol(.),
            missing_percentage = (total_missing / total_values) * 100)

```

# Checking McDonald's omega (Cronbach's alpha alternative)

```{r}
surveys <- c(
  "bat", "ppqr_stress", "ppqr_attitude", "ppqr_resources",
  "copsoq_workload", "copsoq_soc_sup", "copsoq_oppor", "wlbm", "jis",
  "qrp_attitude", "copsoq_infl", "copsoq_meaning", "role_ambiguity",
  "role_conflict", "pay_satisfaction", "tool_satisfaction",
  "general_satisfaction"
)

omega_function <- function(data, survey) {
  items <- data %>%
    dplyr::select(starts_with(survey))

  omega_out <- psych::omega(
    items,
    nfactors = 1,
    plot = FALSE
  )

  data.frame(
    survey = survey,
    omega_total = omega_out$omega.tot
  )
}

omega_results <- data.frame()

for (survey in surveys) {
  omega_result <- omega_function(processed_filtered, survey)
  omega_results <- rbind(omega_results, omega_result)
}

write_csv(omega_results, here::here("analysis/tables/mcdonalds_omega.csv"))
```

# Testing reliability, convergent and discriminant validity
```{r}
library(lavaan)
library(measureQ)

jdr_item_model = '
  WLD =~ copsoq_workload_1 + copsoq_workload_2 + copsoq_workload_3 + copsoq_workload_4
  PPS =~ ppqr_stress_1 + ppqr_stress_2 + ppqr_stress_3 + ppqr_stress_4 + ppqr_stress_5 + ppqr_stress_6
  PPA =~ ppqr_attitude_7 + ppqr_attitude_8 + ppqr_attitude_9 + ppqr_attitude_10 + ppqr_attitude_11 + ppqr_attitude_12
  PPR =~ ppqr_resources_13 + ppqr_resources_14 + ppqr_resources_15 + ppqr_resources_16 + ppqr_resources_17 + ppqr_resources_18
  AMB =~ role_ambiguity_01 + role_ambiguity_02 + role_ambiguity_03 + role_ambiguity_04 + role_ambiguity_05 + role_ambiguity_06
  CON =~ role_conflict_07 + role_conflict_08 + role_conflict_09 + role_conflict_10 + role_conflict_11 + role_conflict_12
  WLB =~ wlbm_1 + wlbm_2 + wlbm_3 + wlbm_4
  OPP =~ copsoq_oppor_1 + copsoq_oppor_2 + copsoq_oppor_3
  MNG =~ copsoq_meaning_1 + copsoq_meaning_2
  INF =~ copsoq_infl_1 + copsoq_infl_2 + copsoq_infl_3 + copsoq_infl_4 + copsoq_infl_5 + copsoq_infl_6
  SOC =~ copsoq_soc_sup_1 + copsoq_soc_sup_2 + copsoq_soc_sup_3 + copsoq_soc_sup_4 + copsoq_soc_sup_5 + copsoq_soc_sup_6
  SEC =~ jis_1 + jis_2 + jis_3 + jis_4
  burnout =~ bat_1 + bat_2 + bat_3 + bat_4
  job_demands =~ WLD + PPA + PPS + AMB + CON + WLB
  job_resources =~ OPP + MNG + INF + SOC + PPR + SEC
  burnout ~ job_demands + job_resources
  '

jdr_final_fit = sem(jdr_item_model, data = processed_filtered_no_missing, std.lv = TRUE, estimator = "MLR")

summary(jdr_final_fit, fit.measures = TRUE, standardized = TRUE, modindices = TRUE, rsquare = TRUE)

manifest_vars <- lavNames(jdr_item_model, type = "ov")

processed_filtered_no_missing <- processed_filtered %>%
  dplyr::select(all_of(manifest_vars)) %>%
  drop_na()

measureQ(jdr_item_model, processed_filtered_no_missing, b.no = 500, HTMT = "TRUE")

jdr_test_model = '
  WLD =~ copsoq_workload_1 + copsoq_workload_2 + copsoq_workload_3 + copsoq_workload_4
  PPS =~ ppqr_stress_1 + ppqr_stress_2 + ppqr_stress_3 + ppqr_stress_4 + ppqr_stress_5 + ppqr_stress_6
  PPA =~ ppqr_attitude_7 + ppqr_attitude_8 + ppqr_attitude_9 + ppqr_attitude_10 + ppqr_attitude_11 + ppqr_attitude_12
  '

measureQ(jdr_test_model, processed_filtered_no_missing, b.no = 500)

Data_A <- read.csv("C:/University/ELTE/Metascience/Psychology of Science/Burnout/pilot_data/researcher_burnout_code/renv/library/windows/R-4.5/x86_64-w64-mingw32/measureQ/extdata/Example_A.csv")

# Specify the model (Model.A) for standardized coefficients
Model.A <- '
              EB =~ EB1 + EB2 + EB3 +EB4 + EB5 + EB6 + EB7 + EB8
              IV =~ IV1 + IV2 + IV3 + IV4 + IV5
              NVG =~ NVG1 + NVG2 + NVG3
              NVA =~ NVA1 + NVA2 + NVA3 + NVA4
             '

# Call the function "measureQ"
measureQ(Model.A, Data_A, b.no=2000)
```

# Aggregating data

```{r}
surveys = c("bat", "ppqr_stress", "ppqr_attitude", "ppqr_resources", "copsoq_workload", "copsoq_soc_sup", "copsoq_oppor", "wlbm", "jis",
            "qrp_attitude", "copsoq_infl", "copsoq_meaning", "role_ambiguity", "role_conflict",
            "pay_satisfaction", "tool_satisfaction", "general_satisfaction", "open_practice")

aggregate_function <- function(data, survey) {
  data %>%
    group_by(anonym_id) %>%
    summarize(
      "{survey}_total" := {
        vars <- across(starts_with(survey))
        n_items <- ncol(vars)
        n_nonmissing <- rowSums(!is.na(vars))

        # Only compute if ≥70% of the items are answered
        complete_enough <- n_nonmissing / n_items >= 0.70

        # Compute row means with NA removal
        out <- rowMeans(vars, na.rm = TRUE)

        # If not complete enough → return NA
        out[!complete_enough] <- NA_real_

        out
      }
    )
}

for (survey in surveys){
  assign(paste0(survey, "_aggregated"), aggregate_function(processed_filtered, survey))
}

aggregated_datasets <- mget(ls(pattern = "_aggregated$"))
aggregated_data <- reduce(aggregated_datasets, left_join, by = "anonym_id")
```

#Joining aggregate data with descriptives

```{r}
aggregated_data = processed_filtered %>% 
  select(anonym_id, gender:lazy_responding) %>% 
  left_join(aggregated_data, by = "anonym_id")
```

#Rename columns

```{r}
aggregated_data_final = aggregated_data %>% 
  rename(burnout = bat_total,
         pub_pressure_stress = ppqr_stress_total,
         pub_pressure_attitude = ppqr_attitude_total,
         pub_pressure_resource = ppqr_resources_total,
         workload = copsoq_workload_total,
         social_support = copsoq_soc_sup_total,
         opportunities = copsoq_oppor_total,
         qrp_attitude = qrp_attitude_total,
         influence = copsoq_infl_total,
         meaning = copsoq_meaning_total,
         role_ambiguity = role_ambiguity_total,
         role_conflict = role_conflict_total,
         pay_satisfaction = pay_satisfaction_total,
         tool_satisfaction = tool_satisfaction_total,
         general_satisfaction = general_satisfaction_total,
         open_practice = open_practice_total,
         job_security = jis_total,
         work_life_balance = wlbm_total)
```

#Save data

```{r}
write.csv(cronbach_results, here::here("analysis/tables/cronbach_alpha.csv"), row.names = FALSE)
write_csv(aggregated_data_final, here::here("data/processed/aggregated_data.csv"))
```
